\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-09-28 paper.qmd
Updated Figure 8

hex_grid_coloured_two_curvy1 <- ggplot() + 
  geom_polygon(
    data = hex_grid_with_counts, 
    aes(x = x, y = y,
      group = hex_poly_id, 
      fill = std_counts), color = "grey70", linewidth=0.2) +
  geom_point(data = umap_two_curvy_scaled,
             aes(x = UMAP1, y = UMAP2),
             alpha = 0.3,
             size = 0.5) +
  scale_fill_viridis_c(direction = -1, 
    na.value = "#ffffff", option = "E") +
  interior_annotation("a", position = c(0.92, 0.93)) + 
  xlim(sc_xlims) + ylim(sc_ylims)

@@

@@

```{r}
#| label: hexbin-regular-two-curvy2
## hexagon binning to have regular hexagons
hb_obj_two_curvy2 <- hex_binning(
  data = umap_two_curvy_scaled, 
  bin1 = 10, 
  r2 = r2)

a1_2 <- calc_bins_y(
  bin1 = 10, 
  r2 = r2
  )$a1

@@

hex_grid_coloured_two_curvy2 <- ggplot() + 
  geom_polygon(
    data = hex_grid_with_counts_two_curvy2, 
    aes(x = x, y = y, 
        group = hex_poly_id, 
        fill = std_counts), 
        color = "grey70", 
        linewidth=0.2) +
  geom_point(data = umap_two_curvy_scaled,
           aes(x = UMAP1, y = UMAP2),
           alpha = 0.3,
           size = 0.5) +
  scale_fill_viridis_c(direction = -1, 
    na.value = "#ffffff", option = "C") +
  xlim(sc_xlims) + ylim(sc_ylims) +
  interior_annotation("b")

@@

```{r}
#| label: hexbin-regular-two-curvy3
## hexagon binning to have regular hexagons
hb_obj_two_curvy3 <- hex_binning(
  data = umap_two_curvy_scaled, 
  bin1 = 13, 
  r2 = r2)

a1_3 <- calc_bins_y(
  bin1 = 13, 
  r2 = r2
)$a1

@@

hex_grid_coloured_two_curvy3 <-  ggplot() + 
  geom_polygon(
    data = hex_grid_with_counts_s_curve3, 
    aes(x = x, y = y, 
        group = hex_poly_id, 
        fill = std_counts), 
        color = "grey70", linewidth=0.2) +
  geom_point(data = umap_two_curvy_scaled,
           aes(x = UMAP1, y = UMAP2),
           alpha = 0.3,
           size = 0.5) +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "D") +
  xlim(sc_xlims) + ylim(sc_ylims) +
  interior_annotation("c")
```

@@

true_pred_df <- predict_emb(
  test_data = true_model_df,
  df_bin_centroids = df_bin_centroids2,
  df_bin = df_bin_two_curvy2,
  type_NLDR = "UMAP"
)

# Compute radius r
a1 <- calc_bins_y(
  bin1 = 13, 
  r2 = r2, 
  q = 0.1)$a1

@@

```{r}
#| echo: false
#| label: fig-bins-two-curvy
#| fig-pos: H
#| fig-cap: "Hexbin density plots of UMAP layout of the S-curve data, using three different bin inputs: (a) $b = 91 \\text{ } (7, \\text{ }13)$, (b) $b = 220 \\text{ } (11, \\text{ }20)$, and (c) $b = 312 \\text{ } (13, \\text{ }24)$. Color indicates standardized counts, dark indicating high count and light indicates low count. At the smallest bin size the data segregates into two separate groups, suggesting this is too many bins. Using the MSE of the model fit in $p\\text{-}D$ helps decide on a useful choice of number of bins."
#| fig-width: 6
#| fig-height: 4

>>>>

hex_grid_coloured_two_curvy1 <- ggplot() + 
  geom_polygon(
    data = hex_grid_with_counts, 
    aes(x = x, y = y,
      group = hex_poly_id, 
      fill = std_counts), color = "grey70", linewidth=0.2) +
  geom_point(data = umap_two_curvy_scaled,
             aes(x = UMAP1, y = UMAP2),
             alpha = 0.3,
             size = 0.5) +
  scale_fill_viridis_c(direction = -1, 
    na.value = "#ffffff", option = "E") +
  interior_annotation("a", sc_ltr_pos) + 
  xlim(c(-0.3, 1.2)) + ylim(c(-0.17, 0.78))

@@

df_bin_two_curvy_temp <- df_bin_two_curvy

@@

```{r}
#| label: hexbin-regular-two-curvy2
## hexagon binning to have regular hexagons
hb_obj_two_curvy2 <- hex_binning(
  data = umap_two_curvy_scaled, 
  bin1 = 12, 
  r2 = r2)

a1_2 <- calc_bins_y(
  bin1 = 12, 
  r2 = r2
  )$a1

@@

hex_grid_coloured_two_curvy2 <- ggplot() + 
  geom_polygon(
    data = hex_grid_with_counts_two_curvy2, 
    aes(x = x, y = y, 
        group = hex_poly_id, 
        fill = std_counts), 
        color = "grey70", 
        linewidth=0.2) +
  geom_point(data = umap_two_curvy_scaled,
           aes(x = UMAP1, y = UMAP2),
           alpha = 0.3,
           size = 0.5) +
  scale_fill_viridis_c(direction = -1, 
    na.value = "#ffffff", option = "C") +
  xlim(c(-0.3, 1.2)) + ylim(c(-0.17, 0.78)) +
  interior_annotation("b", sc_ltr_pos)

@@

```{r}
#| label: hexbin-regular-two-curvy3
## hexagon binning to have regular hexagons
hb_obj_two_curvy3 <- hex_binning(
  data = umap_two_curvy_scaled, 
  bin1 = 14, 
  r2 = r2)

a1_3 <- calc_bins_y(
  bin1 = 14, 
  r2 = r2
)$a1

@@

hex_grid_coloured_two_curvy3 <-  ggplot() + 
  geom_polygon(
    data = hex_grid_with_counts_s_curve3, 
    aes(x = x, y = y, 
        group = hex_poly_id, 
        fill = std_counts), 
        color = "grey70", linewidth=0.2) +
  geom_point(data = umap_two_curvy_scaled,
           aes(x = UMAP1, y = UMAP2),
           alpha = 0.3,
           size = 0.5) +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "D") +
  xlim(c(-0.3, 1.2)) + ylim(c(-0.17, 0.78)) +
  interior_annotation("c", sc_ltr_pos)
```

@@

true_pred_df <- predict_emb(
  test_data = true_model_df,
  df_bin_centroids = df_bin_centroids_two_curvy,
  df_bin = df_bin_two_curvy_temp,
  type_NLDR = "UMAP"
)

# Compute radius r
a1 <- calc_bins_y(
  bin1 = 9, 
  r2 = r2, 
  q = 0.1)$a1

@@

```{r}
#| echo: false
#| label: fig-bins-two-curvy
#| fig-pos: H
#| fig-cap: "Hexbin density plots of UMAP layout of the two-curvy clusters data, using three different bin inputs: (a) $b = 63 \\text{ } (9, \\text{ }7)$, (b) $b = 108 \\text{ } (12, \\text{ }9)$, and (c) $b = 312 \\text{ } (14, \\text{ }10)$. Color indicates standardized counts, dark indicating high count and light indicates low count. Using the MSE of the model fit in $7\\text{-}D$ helps decide on a useful choice of number of bins."
#| fig-width: 9
#| fig-height: 2

### 2024-09-28 paper.tex
Added UMAP first choose initially
Updated Figure 8

\caption{\label{fig-bins-two-curvy}Hexbin density plots of UMAP layout
of the S-curve data, using three different bin inputs: (a)
\(b = 91 \text{ } (7, \text{ }13)\), (b)
\(b = 220 \text{ } (11, \text{ }20)\), and (c)
\(b = 312 \text{ } (13, \text{ }24)\). Color indicates standardized
counts, dark indicating high count and light indicates low count. At the
smallest bin size the data segregates into two separate groups,
suggesting this is too many bins. Using the MSE of the model fit in
\(p\text{-}D\) helps decide on a useful choice of number of bins.}

@@

{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-34-1.pdf}

>>>>

\caption{\label{fig-bins-two-curvy}Hexbin density plots of UMAP layout
of the two-curvy clusters data, using three different bin inputs: (a)
\(b = 63 \text{ } (9, \text{ }7)\), (b)
\(b = 108 \text{ } (12, \text{ }9)\), and (c)
\(b = 312 \text{ } (14, \text{ }10)\). Color indicates standardized
counts, dark indicating high count and light indicates low count. Using
the MSE of the model fit in \(7\text{-}D\) helps decide on a useful
choice of number of bins.}

@@

{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/unnamed-chunk-36-1.pdf}
